---
title: "Practical Machine Learning "
output: html_document
---
## Summary
This R markdown file contains the results of the Peer Assessment of Practical Machine Learning

## Goals of the Assigment 
1) Predicting the manner in which the participants did the exercise while showing the model built and performing cross validation and showing the xpected sample error
2) Use the prediction model to predict 20 different test cases.

## Preparing the Datasets 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r hhi}
library(caret)
library(rpart)
library(knitr)
library(randomForest)
library(ElemStatLearn)
library(ggplot2)
set.seed(696)
```

```{r hi}
trainUrl <-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
trainFile <- "./data/pml-training.csv"
testFile  <- "./data/pml-testing.csv"
if (!file.exists("./data")) {
  dir.create("./data")
}
if (!file.exists(trainFile)) {
  download.file(trainUrl, destfile=trainFile)
}
if (!file.exists(testFile)) {
  download.file(testUrl, destfile=testFile)
}

```

## Read the Data 

```{r pressure}
trainRaw <- read.csv("./data/pml-training.csv",header=T,sep=",",na.strings=c("NA",""))
testRaw <- read.csv("./data/pml-testing.csv",header=T,sep=",",na.strings=c("NA",""))
dim(trainRaw)
dim(testRaw)
```

## Defining the Data Set Parition and Data Cleaning

```{r h}
trainRaw <- trainRaw[,-1] 
inTrain = createDataPartition(trainRaw$classe, p=0.60, list=F)
training = trainRaw[inTrain,]
validating = trainRaw[-inTrain,]
sum((colSums(!is.na(training[,-ncol(training)])) < 0.6*nrow(training)))
Keep <- c((colSums(!is.na(training[,-ncol(training)])) >= 0.6*nrow(training)))
training   <-  training[,Keep]
validating <- validating[,Keep]
```

## Creating the Model

```{r hh}
model <- randomForest(classe~.,data=training)
model
importance(model)
confusionMatrix(predict(model,newdata=validating[,-ncol(validating)]),validating$classe)# model evaluated with confusion matrix
acrcy<-c(as.numeric(predict(model,newdata=validating[,-ncol(validating)])==validating$classe))
acrcy<-sum(acrcy)*100/nrow(validating) # accuracy calculation
```

## Model Testing 

```{r hhh}
testRaw <- testRaw[,-1] 
testRaw <- testRaw[ , Keep] 
testRaw <- testRaw[,-ncol(testRaw)]
# Coerce testing dataset to same class and structure of training dataset 
testing <- rbind(training[100, -59] , testRaw) 
# Apply the ID Row to row.names and 100 for dummy row from testing dataset 
row.names(testing) <- c(100, 1:20)
``` 

## Prediction with the Testing Dataset
```{r hhhh}
predictions <- predict(model,newdata=testing[-1,])
predictions
``` 